Kronos \& Krios: Evidence for accretion of a massive, rocky planetary system in a comoving pair of solar-type stars

### 2017-09-12 paper/ms.tex
final minor edits to abstract and intro

\begin{abstract}
  We report and discuss the discovery of a comoving pair of bright solar-type
  stars, HD~240430 and HD~240429, with a significant difference in their
  chemical abundances.
  The two stars have an estimated 3D separation of $\approx 0.6$~pc
  ($\approx 0.01$~pc projected) at a distance of
  $r\approx 100~\pc$ with nearly identical three-dimensional velocities,
  %($|\Delta \bs{v}| < XX~\kms$ with 95\% confidence),
  as inferred from \gaia\ \tgas\ parallaxes and proper motions, and
  high-precision radial velocity measurements.
  Stellar parameters determined from high-resolution Keck HIRES spectra
  indicate that both stars are $\sim 4$~Gyr old.
  The more metal-rich of the two, HD~240430, shows an enhancement of refractory
  ($\Tcondens>1200$~K) elements by $\approx 0.2$~dex and a marginal
  enhancement of (moderately) volatile elements ($\Tcondens<1200$~K; \elem{C},
  \elem{N}, \elem{O}, \elem{Na}, and \elem{Mn}).
  % TODO: Hmmm...
  This is the largest and the most significant chemical difference found in
  a wide binary pair yet.
  Additionally, HD~240430 shows an anomalously high surface lithium abundance
  ($A(\elem{Li})=2.75$), higher by $0.5$~dex than its companion.
  The proximity in phase-space and ages between the two stars suggests that
  they formed together with the same composition, at odds with the observed
  differences in metallicity and abundance patterns.
  We therefore suggest that the star HD~240430, ``Kronos'', accreted
  \maccreted\ of rocky material after birth, selectively enhancing the
  refractory elements as well as lithium in its surface and convective
  envelope.
\end{abstract}

@@

Binary stars that form from the same birth cloud start with nearly identical
composition.
A differential analysis of the chemical composition of binary stars can reveal
their history through the chemical signatures related to planet formation or
accretion regardless of Galactic chemical evolution.
Stars hosting giant planets on short period orbits have been shown, via
population studies, to form more readily around inherently metal rich stars
(\citealt{Fischer:2005aa,Santos2004}).
However, the post-formation accretion of rocky planets can still alter the
photospheric abundances.
If host stars are polluted after their birth by rocky planetary material with a
high refractory-to-volatile ratio, the convective envelope of the stars may be
enhanced in refractory elements (e.g., \elem{Fe}) compared to their initial
state (e.g., \citealt{Pinsonneault:2001aa}).
Thus, differences in planet formation or accretion in two otherwise identical
stars may imprint differences in chemical abundances that depend on the
condensation temperature (\Tcondens).

High resolution spectroscopic studies of binary star systems hosting at least
one planet (\citealt{2011ApJ...740...76R,2014ApJ...790L..25T,Teske:2013aa,
  Mack:2014aa,Liu:2014aa,Teske:2015aa,Saffe:2015aa,
  Ramirez:2015aa,Biazzo:2015aa,Mack:2016aa,Teske:2016aa,Teske:2016ab})
have yielded varied results:
while some systems appear to have undetectable differences
(see also \citealt{Desidera:2004aa,Gratton:2001aa}), other
studies have reported a \Tcondens-dependent difference in abundance
with higher-\Tcondens\ elements showing larger differences.
A possible explanation for the difference is that
forming more gas giants or rocky planets leads to
an overall or \Tcondens-dependent depletion of high-\Tcondens\ elements in
gas that eventually accretes onto the host star
(\citealt{Ramirez:2015aa,Biazzo:2015aa}).
Alternatively, late time accretion of refractory-rich planetary material can
also produce the trend by enhancing the abundance of high-\Tcondens\ elements
in one of the two stars instead.
The observed differences are $\lesssim 0.1$~dex even in the most dramatic case,
and often closer to $\approx 0.05$~dex, making them challenging to detect even
with a careful analysis of high-resolution, high signal-to-noise ratio spectra,
and differential analyses of two stars that are very similar in their
stellar parameters.
We refer the readers to Appendix~\ref{app:review} for a review of
a handful of individual pairs studied in their detailed chemical abundances
(see also \citealt{2016arXiv161104064M}).

Spectral analysis of polluted white dwarfs (WDs) currently provides the
strongest evidence for accretion of planetary material by a host star
(\citealt{2003ApJ...596..477Z,2010ApJ...722..725Z,2014A&A...566A..34K};
see \citealt{2016NewAR..71....9F} for review).
Because the gravitational settling times of elements heavier than \elem{He} in
atmosphere is much shorter than the WD cooling time
(\citealt{1986ApJS...61..197P}), detection of metals likely indicates the
presence of a reservoir of dusty material around the WD.
Indeed, many of the polluted WDs host a dusty debris disk detected in the
infrared (\citealt{1987Natur.330..138Z,1990ApJ...357..216G,2005ApJ...635L.161R,
  2009ApJ...694..805F,2006ApJ...646..474K}).
Some of the most dramatically polluted WDs show
surface abundances closely matched by rocky planetary material
with, e.g., bulk Earth composition, strongly arguing
that the disk formed from tidally disrupted minor planets
(\citealt{Zuckerman:2007aa,Klein:2010aa}).
Recently, transit signals from small bodies orbiting around a polluted WD
have been detected by \project{Kepler} adding further support to the picture
(\citealt{2015Natur.526..546V}).

Here, we report and discuss the discovery of a comoving pair of G stars,
HD~240430 and HD~240429, with unusual chemical abundance differences that
strongly suggest accretion of rocky planetary material by one of the two stars,
HD~240430.
Throughout the \documentname, we nickname the two stars \bizarreone\
(HD~240430) and \sunanalog\ (HD~240429).
In Greek mythology, Kronos and Krios were sons of Uranos and Gaia.
Kronos notoriously devoured all of his children (except Zeus)
to prevent the prophecy that one day he will be overthrown by them.
We use the following convention for chemical abundances of stars: \elemH{X} is
the log ratio of number density of element \elem{X} to \elem{H} relative to the
solar value.
The absolute abundance of \elem{X} is $A(\elem{X}) = 12 + \log_{10}
(n_\elem{X}/n_\elem{H})$ where $n_\elem{X}$ is the number density of element
\elem{X}.
In \sectionname~\ref{sec:data} we present the astrometric and spectroscopic data
about the two stars relevant to the present discussion.
In \sectionname~\ref{sec:discussion} we discuss possible interpretations of the
abundance difference between the pair.
We summarize our discussions in \sectionname~\ref{sec:summary}.

>>>>

\begin{abstract}
  We report and discuss the discovery of a comoving pair of bright solar-type
  stars, HD~240430 and HD~240429, with a significant difference in their
  chemical abundances.
  The two stars have an estimated 3D separation of $\approx 0.6$~pc
  ($\approx 0.01$~pc projected) at a distance of
  $r\approx 100~\pc$ with nearly identical three-dimensional velocities,
  %($|\Delta \bs{v}| < XX~\kms$ with 95\% confidence),
  as inferred from \gaia\ \tgas\ parallaxes and proper motions, and
  high-precision radial velocity measurements.
  Stellar parameters determined from high-resolution Keck HIRES spectra
  indicate that both stars are $\sim 4$~Gyr old.
  The more metal-rich of the two, HD~240430, shows an enhancement of refractory
  ($\Tcondens>1200$~K) elements by $\approx 0.2$~dex and a marginal
  enhancement of (moderately) volatile elements ($\Tcondens<1200$~K; \elem{C},
  \elem{N}, \elem{O}, \elem{Na}, and \elem{Mn}).
  % TODO: Hmmm...
  This is the largest metallicity difference found in a wide binary pair yet.
  Additionally, HD~240430 shows an anomalously high surface lithium abundance
  ($A(\elem{Li})=2.75$), higher than its companion by $0.5$~dex.
  The proximity in phase-space and ages between the two stars suggests that
  they formed together with the same composition, at odds with the observed
  differences in metallicity and abundance patterns.
  We therefore suggest that the star HD~240430, ``Kronos'', accreted
  \maccreted\ of rocky material after birth, selectively enhancing the
  refractory elements as well as lithium in its surface and convective
  envelope.
\end{abstract}

@@

Binary stars that form from the same birth cloud start with nearly identical
composition.
A differential analysis of the chemical composition of binary stars can reveal
their history through the chemical signatures related to planet formation or
accretion regardless of Galactic chemical evolution.
Giant planets on short period orbits have been shown, via
population studies, to form more readily around inherently metal rich stars
(e.g., \citealt{Fischer:2005aa,Santos2004}).
However, the post-formation accretion of rocky planets can still alter the
photospheric abundances.
If host stars are polluted after their birth by rocky planetary material with a
high refractory-to-volatile ratio, the convective envelope of the stars may be
enhanced in refractory elements (e.g., \elem{Fe}) compared to their initial
state (e.g., \citealt{Pinsonneault:2001aa}).
Thus, differences in planet formation or accretion in two otherwise identical
stars may imprint differences in chemical abundances that depend on the
condensation temperature (\Tcondens).

High resolution spectroscopic studies of binary star systems hosting at least
one planet (\citealt{2011ApJ...740...76R,2014ApJ...790L..25T,Teske:2013aa,
  Mack:2014aa,Liu:2014aa,Teske:2015aa,Saffe:2015aa,
  Ramirez:2015aa,Biazzo:2015aa,Mack:2016aa,Teske:2016aa,Teske:2016ab})
have yielded varied results:
while some systems appear to have undetectable differences
(see also \citealt{Desidera:2004aa,Gratton:2001aa}), other
studies have reported a \Tcondens-dependent difference in abundance
with higher-\Tcondens\ elements showing larger differences.
A possible explanation for the difference is that forming more gas giants or
rocky planets leads to an overall or \Tcondens-dependent depletion of metals in
gas that eventually accretes onto the host star
(\citealt{Ramirez:2015aa,Biazzo:2015aa}).
Alternatively, late time accretion of refractory-rich planetary material can
also produce the trend by enhancing the abundance of high-\Tcondens\ elements
in one of the two stars.
The observed differences are $\lesssim 0.1$~dex even in the most dramatic case,
and often at a level of $\approx 0.05$~dex, making them challenging to detect even
with a careful analysis of high-resolution, high signal-to-noise ratio spectra,
and differential analyses of two stars that are very similar in their
stellar parameters.
We refer the readers to Appendix~\ref{app:review} for a review of
a handful of individual pairs studied in their detailed chemical abundances
(see also \citealt{2016arXiv161104064M}).

Spectral analysis of polluted white dwarfs (WDs) currently provides the
strongest evidence for accretion of planetary material by a host star
(\citealt{2003ApJ...596..477Z,2010ApJ...722..725Z,2014A&A...566A..34K};
see \citealt{2016NewAR..71....9F} for review).
Because the gravitational settling times of elements heavier than \elem{He} in
the WD atmosphere is much shorter than the WD cooling time
(\citealt{1986ApJS...61..197P}), detection of metals likely indicates the
presence of a reservoir of dusty material around the WD.
Indeed, many of the polluted WDs host a dusty debris disk detected in the
infrared (\citealt{1987Natur.330..138Z,1990ApJ...357..216G,2005ApJ...635L.161R,
  2009ApJ...694..805F,2006ApJ...646..474K}).
Some of the most dramatically polluted WDs show
surface abundances closely matched by rocky planetary material
with, e.g., bulk Earth composition, strongly arguing
that the disk formed from tidally disrupted minor planets
(\citealt{Zuckerman:2007aa,Klein:2010aa}).
Recently, transit signals from small bodies orbiting around a polluted WD
have been detected by \project{Kepler} adding further support to the picture
(\citealt{2015Natur.526..546V}).

Here, we report and discuss the discovery of a comoving pair of G stars,
HD~240430 and HD~240429, with unusual chemical abundance differences that
strongly suggest accretion of rocky planetary material by one of the two stars,
HD~240430.
Throughout the \documentname, we nickname the two stars \bizarreone\
(HD~240430) and \sunanalog\ (HD~240429).
In Greek mythology, Kronos and Krios were sons of Uranos and Gaia.
Kronos notoriously devoured all of his children (except Zeus)
to prevent the prophecy that one day he will be overthrown by them.
We use the following convention for chemical abundances of stars:
\elemH{X} is the log ratio of the number density of an element \elem{X} to \elem{H}
relative to the solar value,
$\elemH{X} = \log_{10} ((n_\elem{X}/n_\elem{H})/(n_{\elem{X},~\odot}/n_{\elem{H},~\odot}))$.
The absolute abundance of an element \elem{X} is $A(\elem{X}) = 12 + \log_{10}
(n_\elem{X}/n_\elem{H})$.
In \sectionname~\ref{sec:data} we present the astrometric and spectroscopic data
about the two stars relevant to the present discussion.
In \sectionname~\ref{sec:discussion} we discuss possible interpretations of the
abundance difference between the pair.
We summarize our discussions in \sectionname~\ref{sec:summary}.
